---
title: "Covid19 vaccine study Final 2024"
author: "Matias Lucha"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Males

### Hypothesis and description
$H_0$: COVID921 vaccine is not effective in males.

$H_A$: COVID921 vaccine is effective in males.

### Methods
We will be using the helper file Cat~Cat
Treatment is categorical
Male is categorical

#### Graphical Descriptive Results
```{r}
males=filter(FinalData, FinalData $ sex == "M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males, type="percent")

```

The graphical results show that the perchantage of those who got the placebo were affecred greatly than those who got the vaccine with Covid19. While those with the placebo and compared to the vaccine that were okay were at the same value of about 50%. The vaccine has an affect on those affected as a male according to this chart. 

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows us that those with the placebo were affected by the virus far more than those who were given the vaccine. A comparitive value of 74% with the placebo to a small 26% with the vaccine.

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-values of the Fisher exam and the chi-square test are both 0.00000000000000022. Therefore, the value is lower than 0.05. We therefore reject the null hypothesis, meaning the COVID921 vaccine is effective on males. The odds ratio are 2.82 more in favor of the vaccine over the placebo in males. 

## Females

### Hypothesis and description
$H_0$: Females are independent of effectiveness of the covid921 vaccine.

$H_A$: The effectiveness of the covid921 vaccine is successful female.

### Methods
We will be using the helper file Cat~Cat
Treatment is categorical
Female is categorical

#### Graphical Descriptive Results
```{r}
females=filter(FinalData, FinalData $ sex == "F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

The graphical results show us that females who got the placebo were more affected dramatically by COVID19 than those who received the vaccine. While those who were okay were even statistically being either given the vaccine or placebo. 

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```

Numerical results show us that women who received the vaccine were 35% likely to be infected while those receiving the placebo were 65% likely to get infected. As well as a small difference being more okay having received the vaccine at slightly above 50% whereas the placebo is slightly below 50%. Therefore, the covid921 vaccine is effective in preventing females from contacting COVID19.

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Fisher test and the chi-square test are both below the value of 0.05, therefore we reject the null hypothesis. The chi-square gives us 0.00000001914, and the Fisher test gives us a value of 0.00000001246. Both values are under 0.05, therefore, the COVID921 vaccine is effective on females. With a 1.86 times chance of being affective in an odds ratio. 

## LGBTQ

### Hypothesis and description
$H_0$: Being LGBTQ is independent of effectiveness of the covid921 vaccine.

$H_A$: The effectiveness of the covid921 vaccine is dependent of being LGBTQ.

#### Graphical Descriptive Results
```{r}
barchartGC(~infected + LGBTQ,data=FinalData)
barchartGC(~infected + LGBTQ,data=FinalData, type="percent")

```

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~infected + LGBTQ, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Druggies

### Hypothesis and description

$H_0$: Being a druggie is independent of effectiveness of the covid921 vaccine.

$H_A$: The effectiveness of the covid921 vaccine is dependent of being a druggie

#### Graphical Descriptive Results
```{r}
barchartGC(~infected + DrugUser,data=FinalData)
barchartGC(~infected + DrugUser,data=FinalData, type="percent")

```

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~infected + DrugUser, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

#### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Overall Results and Conclusions